### 2023-09-12 revision/review-response.qmd
plos one revision started

---
format: 
  docx:
    reference-doc: response_styles.docx

execute: 
  echo: false
  warning: false
  message: false
---

```{r echo = F, cache = F, eval = F}
#| eval: false
spelling::spell_check_files(here::here('revision/review-response.qmd'))
```

We sincerely thank the academic editor and two reviewers for providing thoughtful comments on our manuscript.  We provide a point-by-point response to these comments below.

## Academic Editor

Though one of the reviewers recommended a complete rejection, I found that this study could be of benefit to the community especially in the industry 4.0 era. However, I request that you take into consideration the following while revising your manuscript.

1. You should describe explicitly how the package provides a better alternative to all the other existing packages. Where applicable, cite the weaknesses of the existing packages with supporting citations.

__Response__:

2. Make appropriate citations to literature sources. For example, in L26 and L29, you talk about an Act and an EU framework whose sources are not cited. The same applies to L39-40, where you argue out why this communication could be important but without any supporting citation(s).  Similar omissions are evident in the DISCUSSION where you should make appropriate citation of the ‘‘community of practice established’’ and any outcomes (L521-537, and so forth).

__Response__:

3. L87-99 appears to should have been part of the introduction, highlighting what the MassWateR brings onboard.

__Response__:

4. Figure 1 could best be presented as a flow/block diagram.

__Response__:

5. For Table 4, it would be important to provide all the parameters tested as a supplementary file to support the submission. Please ensure that the full details of the algorithms designed are provided, and where there are restrictions, it should otherwise be stated.

__Response__:

6. To improve the chances of manuscript suitability for PLOS ONE, ensure that the criteria listed for manuscript types (https://journals.plos.org/plosone/s/submission-guidelines#loc-methods-software-databases-and-tools) are met. i.e. Utility, Validation and Availability for articles describing new/improved METHODS, SOFTWARE, DATABASES, and TOOLS.

__Response__:

## Reviewer 1

The paper describes a new R package for the processing of water quality information, mainly following US standards, but could be adapted for other regions. It is well written and easy to follow. I have only a few minor comments. Several of the figures were missing from the submission, so I could not check them (Figure 3, and Figure 5 following).

__Response__:

L27 and 29 please cite respective regulatory frameworks

__Response__:

L108 should be „their data in several ways“

__Response__:

L109 should be „outlier checks“

__Response__:

L137 should be function - spelling error

__Response__:

L327 should be „to evaluate“

__Response__:

L334-335 I guess more correct would be to state „to verify that the parameter is below a certain threshold, e.g. below detection limit“. Even if the parameter has a value of zero - the parameter itself is not absent

__Response__:

336-337 same

__Response__:

338-339 More correct would be „to assess the similarity of values, i.e. precision is high. (Precision is a value that is calculated from bot measurements - do not understand how each measurement can have its own precision)

__Response__:

385 Add: „in red in Table 3 show…“

__Response__:

439 spelling „can be“

__Response__:

506 I guess public use is better than public consumption

__Response__:

## Reviewer 2

The availability of reliable monitoring data is the basis of scientific research. This study collected discrete surface water quality data generated by MassWateR software package and analyzed the trend of parameters. The study has some fundamental significance. However, this paper is more like a technical manual than a scientific research paper. Therefore, I would recommend the authors to submit the paper to a technical journal rather than a research-oriented journal.

__Response__:

In addition, the transferability of the developed tool needs to be demonstrated and analyzed, which is directly related to its application value. The authors should give concrete examples in this area;

__Response__:

Finally, the quality of the authors' images is so poor that improving the clarity is very necessary.

__Response__: