### 2023-09-19 supplement.qmd

---
format: 
  docx:
    reference-doc: my_styles.docx
    number-sections: true
title: "*Supplement* MassWateR: Improving quality control, analysis, and sharing of water quality data"
bibliography: refs.bib
csl: plos-one.csl
crossref:
  fig-title: Fig
  fig-prefix: Fig 
author: 
  - Marcus W Beck:
      email: mbeck@tbep.org
      institute: [TBEP]
      correspondence: true
  - Benjamen Wetherill:
      email: bwetherill@acasak.com 
      institute: [ACA]
      correspondence: false
  - Jillian Carr:
      email: Jillian.Carr@umb.edu
      institute: [MASS]
      correspondence: false

institute:
  - TBEP: Tampa Bay Estuary Program, St. Petersburg, Florida, USA
  - ACA: ACASAK Consulting, Boston, Massachusetts USA
  - MASS: Massachusetts Bays National Estuary Partnership, Boston, Massachusetts, USA

filters: 
  - templates/scholarly-metadata.lua
  - templates/author-info-blocks.lua

knitr:
  opts_chunk: 
    fig.path: 'images/'

execute: 
  echo: false
  warning: false
  message: false
  ft.keepnext: false
  ft.align: left
  fig-format: svg
  fig-dpi: 300
---

```{r}
#| include: false
library(MassWateR)
library(dplyr)
library(here)
```

```{r}
#| tbl-cap: "Complete list of parameters and units that can be used with MassWateR.  Parameters can be named using the `Simple Parameter` or the `WQX Parameter` convention.  Only one unit of measure can be used per parameter. WQX: Water Quality Exchange."
#| label: tbl-params
paramsMWR %>% 
  select(`Simple Parameter`, `WQX Parameter`, `Units of measure`) %>% 
  arrange(`Simple Parameter`, .locale = 'en') %>% 
  knitr::kable(.)
```

```{r}
#| tbl-cap: "The sample file included with MassWateR for the data quality objectives for accuracy."
#| label: tbl-acc

accpth <- system.file("extdata/ExampleDQOAccuracy.xlsx", package = "MassWateR")
accdat <- readMWRacc(accpth, warn = F, runchk = F)
knitr::kable(accdat)
```

```{r}
#| tbl-cap: "The sample file included with MassWateR for the data quality objectives for frequency and completeness."
#| label: tbl-frecom
frecompth <- system.file("extdata/ExampleDQOFrequencyCompleteness.xlsx", package = "MassWateR")
frecomdat <- readMWRfrecom(frecompth, warn = F, runchk = F)
knitr::kable(frecomdat)
```