### 2023-10-08 revision2/review-response.qmd
revision2 resubmitted

---
format: 
  docx:
    reference-doc: response_styles.docx
bibliography: ../refs.bib

execute: 
  echo: false
  warning: false
  message: false
---

```{r}
#| echo: false
#| cache: false
#| eval: true
library(MassWateR)

#| results: 'hide'
#| echo: true
# import results data
respth <- system.file("extdata/ExampleResults.xlsx", package = "MassWateR")
resdat <- readMWRresults(respth, warn = F, runchk = F)

# import accuracy data
accpth <- system.file("extdata/ExampleDQOAccuracy.xlsx", package = "MassWateR")
accdat <- readMWRacc(accpth, warn = F, runchk = F)

# import frequency and completeness data
frecompth <- system.file("extdata/ExampleDQOFrequencyCompleteness.xlsx", package = "MassWateR")
frecomdat <- readMWRfrecom(frecompth, warn = F, runchk = F)

# import site data
sitpth <- system.file("extdata/ExampleSites.xlsx", package = "MassWateR")
sitdat <- readMWRsites(sitpth, runchk = F)

# import WQX metadata
wqxpth <- system.file("extdata/ExampleWQX.xlsx", package = "MassWateR")
wqxdat <- readMWRwqx(wqxpth, warn = F, runchk = F)

# a list of input data frames
fsetls <- list(res = resdat, acc = accdat, frecom = frecomdat, sit = sitdat, wqx = wqxdat)
```

**Response**: We again thank the reviewers for evaluating our manuscript. Please see our responses below. Note that reviewer 1 suggested publication without further comment.

## Reviewer 2

The authors have improved over the previous version, but there are still some issues that need to be revised, the main comments are as follows:

1.  It is recommended to add information about the data of the mean values as well as information about the true values in Figures 3 and 6, so that the reader can understand more information;

    **Response**: Figure 3 shows boxplot distributions of selected parameters to aid package users in identifying potential outlier observations. Boxplots are common visualization tools that can quickly demonstrate central tendency and spread of a group of observations [@tukey1977; @mcgill1978], in addition to outliers as intended by the plot in Figure 3. Therefore, we do not consider it necessary to show mean values in this plot. Further, Figure 6 currently shows mean values. The values shown are considered "true" as these are sampled values from *in situ* or laboratory-based methods and not those from modeled or other predictive techniques.

2.  The quality of Figure 4 is very low and it is not even possible to see any content information;

    **Response**: As noted in our responses to comments in the previous revision, the figure quality is substantially reduced in the PDF that is generated by the article submission system. High-resolution figures were included with the submission and can be downloaded by selecting the link that appears at the top-right of each page for each figure.

3.  The dimensions of Figure 8 are recommended to use vertical arrangement; and the scale labeling form is not standardized; the DO color labeling colors of the sites are more distinguishable;

    **Response**: The figures produced by *MassWateR* are all `ggplot` class objects and can be modified using conventional *ggplot2* functions. A vertical arrangement can be achieved as follows, assuming the reviewer means that the legend should appear above and arranged horizontally on the plot.

    ```{r}
    #| fig-height: 5
    #| fig-width: 5
    #| echo: true
    #| out-width: '95%'
    #| fig-align: 'center'
    library(ggplot2)

    anlzMWRmap(fset = fsetls, param = "DO", addwater = "medium") + 
      theme(legend.position = 'top')
    ```

    The online website includes a vignette about modifying *MassWateR* plots with *ggplot2* code. We have added the following text on line 445 to make this clear.

    "Additionally, all plots returned by *MassWateR* are `ggplot` class objects and can be modified accordingly using conventional *ggplot2* functions (see the website [vignette](https://massbays-tech.github.io/MassWateR/articles/modifying.html))."

4.  How do the predictions of water quality parameters in this study compare with other common models?

    **Response**: The functions in *MassWateR* rely on data collected *in situ* and there are no aspects related to predictive modeling included in the package. The primary goals of *MassWateR* are to improve how environmental professionals perform quality control, analysis, and sharing of monitoring data. As such, the intent is to support the use of data for environmental decision-making.  No predictions are provided by the package.

5.  Some recent deep learning models for water quality prediction should be mentioned, e.g., 10.1016/j.ejrh.2023.101331; 10.1007/s11783-023-1688-y; 10.1016/j.ecolind.2023.109882; 10.1016/j.jhydrol.2023. 129649.

    **Response**: Please see our response to the previous comment.

# References {.unnumbered}